---
title: "plotting_markdown"
output: html_document
date: "2024-01-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mini project 2: Causal relationship between ADHD stimulant medication and 


##initialise the document

```{r intro}
##imports
#install.packages("extrafont")
library(extrafont)
library(tidyverse)
library(ggdag)
library(dagitty)
library(ggplot2) 
library(texreg)
library(cowplot)
library(cobalt)
library(geepack)




##LOAD IN DATASET CRETEAD IN PYTHON USING 'raw_data_manipulation.ipynb or .py
for_regression_130224 <- read_csv("~/Documents/School/phd/mini-projects/Kelleher_block_2/code/data/processed/for_regression_130224.csv")

data_subject_subset <- subset(for_regression_130224, subset_subjects == 1)


```


##plot DAG 


```{r plot dag}

##we start by plotting DAG 
model_from_dagitty <- dagitty('dag {

"Mental health score" [pos="-1.25,-0.803"]
"Pyschotic experiences" [outcome,pos="-0.5,-0.297"]
"Sex" [confounder,pos="-0.75,-0.802"]
"SEP" [pos="-0.250,-0.803"]
"Stimulant medication" [exposure,pos="-1.5,-0.293"]
"Urban Rural" [pos="0.25,-0.803"]
ADHD [adjusted,pos="-1.75,-0.803"]

"Mental health score" -> "Pyschotic experiences"
"Mental health score" -> "Stimulant medication"
"Sex" -> "Pyschotic experiences"
"Sex" -> "Stimulant medication"
"SEP" -> "Pyschotic experiences"
"SEP" -> "Stimulant medication"
"Stimulant medication" -> "Pyschotic experiences"
ADHD -> "Pyschotic experiences"
ADHD -> "Stimulant medication"\
ADHD -> "Mental health score"

"Urban Rural" -> "Pyschotic experiences"
"Urban Rural" -> "Stimulant medication"


}
')

ggdag_status(model_from_dagitty,text = FALSE, use_labels = 'name') +
  theme_dag()


ggdag_adjustment_set(model_from_dagitty,use_labels = 'name')
```
##prep the data

1. take the subjects that dont have nans in any of the latent factors
2. make new categorical variable of treatment group membership (at least of one year of stimulant use, no baseline -> done in python earlier)

```{r prep}
data_subject_subset_1<- data_subject_subset[
  !is.na(data_subject_subset$ADHD_baseline_year_1_arm_1) & 
    !is.na(data_subject_subset$PE_3_year_follow_up_y_arm_1) &
    !is.na(data_subject_subset$PE_baseline_year_1_arm_1) &
  !is.na(data_subject_subset$CI_average) &
  !is.na(data_subject_subset$EDU_HIGER_AVERAGE) &
    !is.na(data_subject_subset$UR_0_baseline_year_1_arm_1)
  , ]

data_subject_subset_1$TreatmentGroup <- ifelse(data_subject_subset_1$meds_nbl_and_y1_or_y2_and_y3 == 1, TRUE, FALSE)

data_subject_subset_1 %>% 
  select(Subject,meds_nbl_and_y1_or_y2_and_y3,TreatmentGroup) %>%
  head(100) #check it works, where meds 0 we should have FALSE and one ie



```



##regression model

1. we start by plotting the simple group identity

2. we build a regression model that 

```{r cowplot}

means_PE <- aggregate(PE_3_year_follow_up_y_arm_1 ~  TreatmentGroup, data_subject_subset_1, mean)
means_ADHD <- aggregate(ADHD_baseline_year_1_arm_1 ~  TreatmentGroup, data_subject_subset_1, mean)




ggplot(data_subject_subset_1, aes(x = TreatmentGroup, y = PE_3_year_follow_up_y_arm_1),
       ylim =c(0,3)) +
geom_violin() +
stat_summary(fun=mean, colour="darkred", geom="point", 
               shape=18, size=3, show.legend=FALSE  ) + 
geom_text(data = means_PE, aes(label = PE_3_year_follow_up_y_arm_1, y = PE_3_year_follow_up_y_arm_1 - 1)) +
cowplot::theme_cowplot(font_size = 11) +
  coord_cartesian(ylim =c(0,20))  



ggplot(data_subject_subset_1, aes(x = TreatmentGroup,, y = ADHD_baseline_year_1_arm_1),
       ylim =c(0,3)) +
stat_summary(fun=mean, colour="darkred", geom="point", 
               shape=18, size=3, show.legend=FALSE) + 
geom_text(data = means_ADHD, aes(label = ADHD_baseline_year_1_arm_1, y = ADHD_baseline_year_1_arm_1 + 12.08)) +
geom_violin() +
cowplot::theme_cowplot(font_size = 11)
#coord_cartesian(ylim =c(0,10))

##regression models

model_naive <- lm(PE_3_year_follow_up_y_arm_1 ~ TreatmentGroup, data = data_subject_subset_1)
#texreg::screenreg(model_naive)
summary(model_naive)
confint(model_naive)
boxplot(PE_3_year_follow_up_y_arm_1 ~ TreatmentGroup, data = data_subject_subset_1,ylim = c(0,20))
abline(model_naive, colour = 'blue' )


model_compelx <- lm(PE_3_year_follow_up_y_arm_1 ~ Sex_at_birth + ADHD_baseline_year_1_arm_1 + TreatmentGroup + PE_baseline_year_1_arm_1, data = data_subject_subset_1)
#texreg::screenreg(model_naive)
summary(model_compelx)


```



## matching model

Here we use the subject matching as our model.

```{r pressure, echo=FALSE}

library(MatchIt)

matched <- matchit(TreatmentGroup ~ Sex_at_birth + ADHD_baseline_year_1_arm_1 + PE_baseline_year_1_arm_1 + CI_average + EDU_HIGER_AVERAGE + UR_0_baseline_year_1_arm_1, data = data_subject_subset_1,
                    methods = 'full', distance = 'glm', replace = FALSE,ratio = 1)
matched

data_subject_subset_1_matched <- match.data(matched)

model_matched <- lm(PE_3_year_follow_up_y_arm_1 ~ TreatmentGroup, data = data_subject_subset_1_matched,
weights = weights)
texreg::screenreg(model_matched)

```
```{r ipw}

model_treatment <- glm(TreatmentGroup ~ Sex_at_birth + ADHD_baseline_year_1_arm_1 + PE_baseline_year_1_arm_1 + CI_average + EDU_HIGER_AVERAGE + UR_0_baseline_year_1_arm_1 + anxdisord_baseline_year_1_arm_1 + depress_baseline_year_1_arm_1, data = data_subject_subset_1,
family = binomial(link = "logit"))
edu_age_propensities <- broom::augment_columns(model_treatment, data_subject_subset_1,
type.predict = "response") %>%
rename(propensity = .fitted)

edu_age_propensities %>%
select(Subject, TreatmentGroup, Sex_at_birth, ADHD_baseline_year_1_arm_1, PE_baseline_year_1_arm_1,CI_average,propensity) %>%
arrange(desc(propensity)) %>%
head(100)


df_ipws <- edu_age_propensities %>%
  mutate(ipw = (TreatmentGroup / propensity) +
           (1 - TreatmentGroup)/(1 - propensity))



df_ipws %>%
select(Subject, TreatmentGroup, Sex_at_birth, ADHD_baseline_year_1_arm_1, PE_baseline_year_1_arm_1,CI_average,anxdisord_baseline_year_1_arm_1, EDU_HIGER_AVERAGE, UR_0_baseline_year_1_arm_1,ipw) %>%
arrange(desc(ipw)) %>%
head(100)

ggplot(df_ipws, aes(x = ADHD_baseline_year_1_arm_1, y = EDU_HIGER_AVERAGE,
color = TreatmentGroup, size = ipw)) +
geom_point()




model_ipw <- lm(PE_3_year_follow_up_y_arm_1 ~ TreatmentGroup, data = df_ipws,
weights = ipw)
texreg::screenreg(model_ipw)


```
```{r}
df_ipws$weighted <- df_ipws$PE_3_year_follow_up_y_arm_1 * df_ipws$ipw

means_PE <- aggregate(PE_3_year_follow_up_y_arm_1 ~  TreatmentGroup, data_subject_subset_1, mean)


ggplot(df_ipws, aes(x = TreatmentGroup, y = weighted),
       ylim =c(0,3)) +
geom_violin() +
#stat_summary(fun=mean, colour="darkred", geom="point", 
               #shape=18, size=3, show.legend=FALSE  ) + 
#geom_text(data = means_PE, aes(label = PE_3_year_follow_up_y_arm_1, y = PE_3_year_follow_up_y_arm_1 - 1)) +
cowplot::theme_cowplot(font_size = 11) #+
  #coord_cartesian(ylim = c(0,20))
```

```{r ipw using hernan robins example}
library("geepack")


for_regression_170224 <- read_csv("~/Documents/School/phd/mini-projects/Kelleher_block_2/code/data/processed/for_regression_190224_raw_scores.csv")

data_subject_subset_1702 <- subset(for_regression_170224, subset_subjects == 1)


#data_subject_subset_1702 
data_subject_subset_1702 <- data_subject_subset_1702[
  !is.na(data_subject_subset_1702$ADHD_baseline_year_1_arm_1) & 
    !is.na(data_subject_subset_1702$PE_3_year_follow_up_y_arm_1) &
    #!is.na(data_subject_subset_1702$PE_baseline_year_1_arm_1) &
  !is.na(data_subject_subset_1702$CI_average) &
  !is.na(data_subject_subset_1702$EDU_HIGER_AVERAGE) &
    !is.na(data_subject_subset_1702$UR_0_baseline_year_1_arm_1) &
    !is.na(data_subject_subset_1702$anxdisord_baseline_year_1_arm_1)&
    !is.na(data_subject_subset_1702$depress_baseline_year_1_arm_1)
  , ]

data_subject_subset_1702$TreatmentGroup <- ifelse(data_subject_subset_1702$meds_nbl_and_y1_or_y2_and_y3 == 1, TRUE, FALSE)


denom_fit <- glm(
  TreatmentGroup ~ Sex_at_birth + ADHD_baseline_year_1_arm_1 + #PE_baseline_year_1_arm_1 +
    CI_average + EDU_HIGER_AVERAGE + UR_0_baseline_year_1_arm_1 + 
    anxdisord_baseline_year_1_arm_1 + depress_baseline_year_1_arm_1
  ,
  family = binomial(),
  data = data_subject_subset_1702
)


summary(denom_fit)


numer_fit <- glm(
  TreatmentGroup ~ 1, 
  family = binomial(),
  data = data_subject_subset_1702
)
summary(numer_fit)


pd.qsmk <- predict(denom_fit, type = "response")

pn.qsmk <- predict(numer_fit, type = "response")

data_subject_subset_1702$sw <-
  ifelse(data_subject_subset_1702$TreatmentGroup == 0, ((1 - pn.qsmk) / (1 - pd.qsmk)),
         (pn.qsmk / pd.qsmk))


summary(data_subject_subset_1702$sw)


data_subject_subset_1702$Row_ID <- row.names(data_subject_subset_1702)


msm.sw <- geeglm(
  PE_3_year_follow_up_y_arm_1 ~ TreatmentGroup,
  data = data_subject_subset_1702,
  weights = sw,
  id = Row_ID,
  corstr = "independence"
)


summary(msm.sw)



```
```{r}

ggplot(data_subject_subset_1702, aes(x = ADHD_baseline_year_1_arm_1, y = CI_average,
color = TreatmentGroup, size = sw)) +
geom_point()


xtabs(data_subject_subset_1702$sw ~ data_subject_subset_1702$Sex_at_birth + data_subject_subset_1702$TreatmentGroup)


ggplot(data_subject_subset_1702, aes(x = sw, fill = TreatmentGroup)) +
  geom_density(alpha = 0.5) +
  labs(title = "Comparison of Two Distributions",
       x = "weight",
       y = "Density") +
  scale_fill_manual(values = c("blue", "red")) 


```
```{r}
data_subject_subset_1702 %>%
select(Subject, TreatmentGroup, Sex_at_birth, ADHD_baseline_year_1_arm_1, PE_baseline_year_1_arm_1,CI_average,anxdisord_baseline_year_1_arm_1, EDU_HIGER_AVERAGE, UR_0_baseline_year_1_arm_1,sw) %>%
arrange(desc(sw)) %>%
head(100)

data_frame_reduct = data_subject_subset_1702 %>%
select(Subject, TreatmentGroup, Sex_at_birth, ADHD_baseline_year_1_arm_1, PE_baseline_year_1_arm_1,CI_average,anxdisord_baseline_year_1_arm_1, EDU_HIGER_AVERAGE, UR_0_baseline_year_1_arm_1,sw)

means <- aggregate(ADHD_baseline_year_1_arm_1 ~ TreatmentGroup, data = data_subject_subset_1702, FUN = mean)

means_weigted <- aggregate((sw*ADHD_baseline_year_1_arm_1) ~ TreatmentGroup, data = data_subject_subset_1702, FUN = mean)



print(means)
print(means_weigted)

ggplot(data_subject_subset_1702, aes(x = sw*ADHD_baseline_year_1_arm_1, fill = TreatmentGroup)) +
  geom_density(alpha = 0.5) +
        stat_summary(aes(xintercept = ..x.., y = 0,colour = TreatmentGroup), fun = mean, geom = "vline", orientation = "y") +

  labs(title = "Comparison of Two Distributions",
       x = "weight * ADHD",
       y = "Density") +
  scale_fill_manual(values = c("blue", "red")) 

ggplot(data_subject_subset_1702, aes(x = ADHD_baseline_year_1_arm_1, fill = TreatmentGroup)) +
  geom_density(alpha = 0.5) +
    stat_summary(aes(xintercept = ..x.., y = 0,colour = TreatmentGroup), fun = mean, geom = "vline", orientation = "y") +
  labs(title = "Comparison of unbalanced Distributions",
       x = "ADHD",
       y = "Density") +
  scale_fill_manual(values = c("blue", "red")) 

ggplot(data_subject_subset_1702, aes(y = sw*ADHD_baseline_year_1_arm_1, x = TreatmentGroup)) +
  geom_boxplot(alpha = 0.5) +
      stat_summary(aes(xintercept = ..x.., y = 0,colour = TreatmentGroup), fun = mean, geom = "vline", orientation = "y") +
  labs(title = "Comparison of balanced Distributions",
       x = "Treatment Group",
       y = "weighted ADHD_baseline ") +
  scale_fill_manual(values = c("blue", "red")) 


ggplot(data_subject_subset_1702, aes(y = ADHD_baseline_year_1_arm_1, x = TreatmentGroup)) +
  geom_boxplot(alpha = 0.5) +
  labs(title = "Comparison of Two Distributions",
       x = "Treatment Group",
       y = "unweighted ADHD_baseline ") +
  scale_fill_manual(values = c("blue", "red")) 




ggplot(data_subject_subset_1702, aes(y = sw*EDU_HIGER_AVERAGE, x = TreatmentGroup)) +
  geom_boxplot(alpha = 0.5) +
  labs(title = "Comparison of Two Distributions",
       x = "Treatment Group",
       y = "weighted edu ") +
  scale_fill_manual(values = c("blue", "red")) 


ggplot(data_subject_subset_1702, aes(y = EDU_HIGER_AVERAGE, x = TreatmentGroup)) +
  geom_boxplot(alpha = 0.5) +
  labs(title = "Comparison of Two Distributions",
       x = "Treatment Group",
       y = "edu ") +
  scale_fill_manual(values = c("blue", "red"))

ggplot(data_subject_subset_1702, aes(y = sw*UR_0_baseline_year_1_arm_1, x = TreatmentGroup)) +
  geom_boxplot(alpha = 0.5) +
  labs(title = "Comparison of Two Distributions",
       x = "Treatment Group",
       y = "weighted edu ") +
  scale_fill_manual(values = c("blue", "red")) 


ggplot(data_subject_subset_1702, aes(y = UR_0_baseline_year_1_arm_1, x = TreatmentGroup)) +
  geom_boxplot(alpha = 0.5) +
  labs(title = "Comparison of Two Distributions",
       x = "Treatment Group",
       y = "edu ") +
  scale_fill_manual(values = c("blue", "red")) 


ggplot(data_subject_subset_1702, aes(x = sw, fill = TreatmentGroup)) +
  geom_histogram(alpha = 0.5) +
        stat_summary(aes(xintercept = ..x.., y = 0,colour = TreatmentGroup), fun = mean, geom = "vline", orientation = "y") +

  labs(title = "Comparison of Two Distributions",
       x = "weight",
       y = "Density") +
  scale_fill_manual(values = c("blue", "red")) 




```
```{r}
##weighted sum 
print(sum(data_subject_subset_1702$sw * data_subject_subset_1702$ADHD_baseline_year_1_arm_1) / sum(data_subject_subset_1702$sw))

table(data_subject_subset_1702$TreatmentGroup)
```
```{r}


custom_labels <- c(Sex_at_birth = "Sex at Birth",
                   ADHD_baseline_year_1_arm_1 = "ADHD at Baseline",
                   EDU_HIGER_AVERAGE = "Average Education Level",
                   UR_0_baseline_year_1_arm_1 = "UR at Baseline",
                   PE_baseline_year_1_arm_1 = "PE at Baseline",
                   CI_average = "Average CI"
)
love.plot(TreatmentGroup ~ Sex_at_birth + ADHD_baseline_year_1_arm_1 + EDU_HIGER_AVERAGE + UR_0_baseline_year_1_arm_1 + PE_baseline_year_1_arm_1 + CI_average,  
          data = data_subject_subset_1702, 
          weights = W.out$w,
          abs = TRUE,
          line = FALSE, 
          thresholds = c(m = .1),
          colors = c("red", "blue"),
          method = "weighting", estimand = "ATE",
          var.names = custom_labels)



bal.tab(W.out, threshold=0.1)
```
```{r}
(sum((data_subject_subset_1702$sw))^2)/(sum((data_subject_subset_1702$sw)^2))


(sum((df_ipws$ipw))^2)/(sum((df_ipws$ipw)^2))


library(WeightIt)
W.out <- weightit(TreatmentGroup ~ Sex_at_birth + ADHD_baseline_year_1_arm_1 + PE_baseline_year_1_arm_1 +
    CI_average + EDU_HIGER_AVERAGE + UR_0_baseline_year_1_arm_1 + 
    anxdisord_baseline_year_1_arm_1 + depress_baseline_year_1_arm_1,
                  data = data_subject_subset_1702, estimand = "ATE", method = "gbm", stabilize=FALSE)

summary(W.out)



msm.sw <- geeglm(
  PE_3_year_follow_up_y_arm_1 ~ TreatmentGroup,
  data = data_subject_subset_1702,
  weights = W.out$weights,
  id = Row_ID,
  corstr = "independence"
)


summary(msm.sw)



(sum((W.out$weights))^2)/(sum((W.out$weights)^2))

```
```{r final model}
library(WeightIt)
library(gbm)

for_regression_200224 <- read_csv("~/Documents/School/phd/mini-projects/Kelleher_block_2/code/data/processed/for_regression_270224_raw_scores.csv") ##read the csv

data_subject_subset_2002 <- subset(for_regression_200224, subset_subjects == 1) ##from the whole data choose the subjects who are in either control group (no stimulants) or treatment (at least of year of stimulant use)

##define the treatment group
data_subject_subset_2002$TreatmentGroup <- ifelse(data_subject_subset_2002$meds_nbl_and_y1_or_y2_and_y3 == 1, TRUE, FALSE) ##make the treatment group TRUE FALSE COLUMN


data_subject_subset_2002$TreatmentGroup_level <- factor(data_subject_subset_2002$TreatmentGroup, levels = unique(data_subject_subset_2002$TreatmentGroup))


print('here')


summary(data_subject_subset_2002$TreatmentGroup) 
# Vector of column names to check for NaN values

 ##Define the columns we are going to use from the whole dataset (see the DAG)
columns_to_check <- c( 
  "Sex_at_birth",
  #"Gender_Conformity",
  "ADHD_baseline_year_1_arm_1",
  #"ADI_0_baseline_year_1_arm_1",
  "EDU_HIGER_AVERAGE_squashed",
  "depress_baseline_year_1_arm_1",
  "anxdisord_baseline_year_1_arm_1",
  "somaticpr_baseline_year_1_arm_1",
  "latent_factor_ss_perinatal_baseline_year_1_arm_1",
  "race_ethnicity_written",
  "PE_baseline_year_1_arm_1",
  "CI_squashed",
  'UR_0_baseline_year_1_arm_1',
  #'EDU_HIGER_AVERAGE',
  
  ##treatment group
  'TreatmentGroup',
  'TreatmentGroup_level',
  'meds_nbl_and_y1_or_y2_and_y3',
  
  ##this is the dependent variable
  "PE_3_year_follow_up_y_arm_1"
)

cofactros = columns_to_check[!columns_to_check %in% "PE_3_year_follow_up_y_arm_1"]

#colSums(is.na(data_subject_subset_2002[,cofactros]))


# Indexing operation to remove rows with NaN values in specified columns
data_subject_subset_2002 <- data_subject_subset_2002[complete.cases(data_subject_subset_2002[, columns_to_check]), ]


#as.factor(data_subject_subset_2002$race_ethnicity_written)

W.out <- weightit(meds_nbl_and_y1_or_y2_and_y3~ . + CI_squashed * ADHD_baseline_year_1_arm_1 + anxdisord_baseline_year_1_arm_1 * ADHD_baseline_year_1_arm_1,
                  data = data_subject_subset_2002[,cofactros], estimand = "ATT", method = "gbm", stabilize=TRUE,thresholds = c(m = .1),trim.at = 0 ,focal = 1.0) 


##gbm = Generalized Boosted Models


```

```{r}
 your_treatment_variable = (data_subject_subset_2002$TreatmentGroup)

levels(your_treatment_variable)
factor(your_treatment_variable, levels = unique(your_treatment_variable))

```


```{r}
library(cobalt)
library(geepack)
summary(W.out)
love.plot(TreatmentGroup ~ ., 
          data = data_subject_subset_2002[,cofactros],
          weights = W.out$w,
          abs = FALSE,
          line = FALSE, 
          thresholds = c(m = .1),
          colors = c("red", "blue"),
          method = "weighting", estimand = "ATE")


data_subject_subset_2002$Row_ID <- row.names(data_subject_subset_2002)

msm.sw <- geeglm(
  PE_3_year_follow_up_y_arm_1 ~ TreatmentGroup,
  data = data_subject_subset_2002,
  weights = W.out$weights,
  id = Row_ID,
  corstr = "independence"
)

summary(msm.sw)

```
```{r}
summary(W.out)
```


